hbm_paper_results

### 2024-05-17 paper_main_results.Rmd
cleaning up

Rmd to show and get feedback on planned for results & figures for the combatLS paper. Currently thinking I'll have centile score results be the focus of the paper and keep z-scores for the supplement.

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x)
#Sys.setenv("MC_CORES"=10L)
options("mc.cores")
```
```{r}
source("./cubic_scripts/R_scripts/stats_tests.R")
# devtools::source_url("https://raw.githubusercontent.com/BGDlab/combat-biovar/main/cubic_scripts/R_scripts/stats_tests.R")
```

@@

# Main Analyses

## Load Data

@@

**QC: find ppts with centiles at 1 or 0**

```{r}
combined_preds %>% 
  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0))) %>%
  dplyr::select(participant) %>%
  unique()

combined_preds %>% 
  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
  dplyr::select(participant) %>%
  unique()
```

saving out csv forces some centile scores to round up to 1, BUT it appears z scores are still being calculated correctly so just going to proceed!

@@

Get differences in male vs female predicted variances at mean sample age

```{r}
pred.csvs <- list.files(path = paste0(data_path, "/ukb_basic"), pattern = "_variance.csv", full.names = TRUE)
combined_var <- data.frame()

  # Add a "Source_File" column with the file name
  data <- data %>%
    mutate(Source_File = as.factor(basename(file))) %>%
    mutate(dataset = sub("_data_variance.csv", "", Source_File),
           sex_standard = sex_effect/f.var)

  # Bind the data to the combined dataframe
  combined_var <- bind_rows(combined_var, data, .id = "File_ID")
}

#get summary df for p vals, plotting
sigma.sex.df <- fread("ukb_basic/gamlss_summary.csv", stringsAsFactors = TRUE) %>%
  dplyr::select(-V1) %>%
  mutate(dataset = gsub("-data", "", dataset),
         dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset)) %>% #clean up dataset naming
  mutate(dataset = factor(dataset, levels = unique(dataset), ordered = FALSE)) %>%
  mutate(dataset = relevel(dataset, ref= "ukb_CN")) %>%
  dplyr::filter(parameter == "sigma" & term == "sexMale") %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, sig.sum_bf.corr)

sigma.sex.df <- left_join(sigma.sex.df, dk.parc, by="label")

plot.info <- sigma.sex.df %>%
  dplyr::select(mod_name, pheno, dataset, pheno_cat, label, hemi, region, sig.sum_bf.corr) %>%
  unique() %>%
  mutate(dataset = gsub("-", "_", dataset),
         dataset = gsub("_simsites", "", dataset))

combined_var <- inner_join(combined_var, plot.info)

#drop refA sites, not using for these analyses
combined_var <- combined_var %>%
  dplyr::filter(!grepl('_refA', dataset)) #drop refA levels
```

## SCALE EFFECTS OF SEX

```{r}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
#var plot
var.plot.df <- combined_var %>%
  dplyr::filter(sig.sum_bf.corr == TRUE & pheno_cat != "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset, pheno_cat)

tp <- unique(var.plot.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

var.plot <- ggplot(var.plot.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = dk, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_standard)) +
  #scale_fill_gradient2() +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
#var.plot

#ggsave("figs/male_variance_ohbm.jpeg", var.plot, width=10, height=10)
```

```{r}
#plot global
var.plot.global.df <- combined_var %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  dplyr::filter(!grepl('_refA', dataset)) %>% #drop refA levels
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  mutate(tissue_class = as.factor(case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA)),
         sex_standard = ifelse(sig.sum_bf.corr == TRUE, sex_standard, NA)) %>% #drop NS values
  group_by(dataset)

tp <- unique(var.plot.global.df[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="ukb_CN")
tp$sex_standard <- 1

var.plot.global <- ggplot(var.plot.global.df) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = sex_standard),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
#var.plot.global
```

Combine plots
```{r}
p <- ggarrange(
  var.plot.global, var.plot,
  ncol=2, nrow=1, 
  widths = c(1, 3.1),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "bottom"
  )

p_annotate <- annotate_figure(p, top = text_grob("Figure 1B. Standardized effect of male sex on feature variance", 
               color = "black", size = 14, hjust=.65))
p_annotate
ggsave("figs/male_variance_ohbm_big.jpeg", p_annotate, width=7, height=8)
```
Look at relative estimates for sex in sigma across features in raw data

```{r}
gamlss.sum.df <- fread(paste0(data_path, "/ukb_basic/gamlss_summary.csv"))
gamlss.sum.df.raw <- gamlss.sum.df %>%
  dplyr::filter(dataset=="ukb-CN-data-simsites")

gamlss.sum.df.raw.sexsig <- gamlss.sum.df.raw %>%
  dplyr::filter(parameter =="sigma" & term == "sexMale") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      est_over_moment = estimate/moment.mean)

gamlss.sum.df.raw.sexsig %>%
  ggplot(aes(y=estimate, x=pheno_cat, fill=pheno_cat)) +
  geom_boxplot() +
  # geom_hline(yintercept = mean(gamlss.sum.df.raw.sexsig$t_stat), linetype = 2) +
  stat_compare_means(method = "anova", label.y = 0.07)+    
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")

gamlss.sum.df.raw.sexsig %>%
  ggplot() +
  geom_histogram(aes(x=est_over_moment, fill=pheno_cat, color=pheno_cat), alpha=0.5, position="identity")

gamlss.sum.df.raw.sexsig %>%
  ggplot(aes(y=t_stat, x=pheno_cat, fill=pheno_cat)) +
  geom_boxplot() +
  # geom_hline(yintercept = mean(gamlss.sum.df.raw.sexsig$t_stat), linetype = 2) +
  stat_compare_means(method = "anova", label.y = -7)+    
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = ".all.")
```

## CENTILES

Do centile scores change based on combat version?
Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)

@@

```{r, eval=FALSE}
full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
saveRDS(full_cent_res, "./derivatives/ukb_basic/featurewise_cent_t_tests_all_out.RDS")
```

@@

```{r, eval=FALSE}
full_cent_res%>%
  dplyr::filter(comp =="cf.gamlss_cf" & bigger_group == "cf.gamlss" & sig_fdr ==TRUE)

diffs.df %>%
  dplyr::filter(dataset=="cf.gamlss" | dataset =="cf") %>%
  ggplot() +
  geom_boxplot(aes(y=abs.diff_rh_CT_parstriangularis, fill=dataset))

diffs.df %>%
  dplyr::filter(dataset=="cf.gamlss" | dataset =="cf") %>%
  ggplot() +
  geom_boxplot(aes(y=diff_rh_CT_parstriangularis, fill=dataset))
```

@@

get stat results for global phenos:
```{r, message=FALSE}
full_cent_res %>%
  dplyr::filter(pheno %in% vol_list_global) %>%
  dplyr::filter(!grepl('_refA', comp))

full_cent_res %>%
  dplyr::filter(bigger_group == "cf.gamlss")%>%
  dplyr::filter(!grepl('_refA', comp))
```

all significant

```{r}
global_abs.diffs <- diffs.df %>%
  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", vol_list_global)) %>%
  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")

@@

```{r}
gam.gamlss.diffs <- diffs.df %>%
  dplyr::filter(dataset=="cf.gam" | dataset=="cf.gamlss")

gam_better <- gam.gamlss.diffs %>%
  group_by(dataset) %>%
  summarize_at(pheno_abs.diff.list, median) %>%
  ungroup() %>%
  pivot_longer(cols = all_of(pheno_abs.diff.list), names_to="pheno") %>%
  pivot_wider(names_from = dataset, values_from=value, names_glue = "med_{dataset}") %>%
  mutate(gam_minus_gamlss = (med_cf.gam - med_cf.gamlss),
         pheno = sub("abs.diff_", "", pheno)) %>%
  ungroup() %>%
  dplyr::filter(gam_minus_gamlss < 0)

ns_gam_gamlss_comp <- full_cent_res %>%
  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE)

cfgam.better <- intersect(gam_better$pheno, ns_gam_gamlss_comp$pheno)
```
```{r, eval=FALSE}
cfgam.better.ct <- gsub("_CT", "", cfgam.better)
cfgam.better.ct_plot <- dk.parc %>%
  dplyr::mutate(gam_better = ifelse(label %in% cfgam.better.ct, 1, 0))

ggplot(cfgam.better.ct_plot) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
            aes(fill=as.factor(gam_better)))
```

```{r, eval=FALSE}
for (pheno in pheno_abs.diff.list){
  p <- ggplot(gam.gamlss.diffs, aes(y = gam.gamlss.diffs[[pheno]], fill= dataset)) +
      geom_boxplot() +
    ggtitle(pheno)
print(p)
}
```

@@

### Sex-Biases

Playing with different ways to visualize each subject's average centile error (i.e., raw data centile - combatted data centile) and magnitude centile error (absolute value)

```{r, eval=FALSE}
subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_boxplot(aes(x=sex, y=mean_cent_abs.diff, fill=sex)) +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  facet_wrap(~dataset) +
  theme(legend.position="none") +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat configurations")

subj_sum.df %>%
  mutate(mean_cent_diff = mean_cent_diff*100) %>% #for visuals
  ggplot(aes(x=dataset, y=mean_cent_diff, fill=sex)) +
  geom_violin(trim=FALSE, scale="width") +
  labs(y="Centile Error", x = "ComBat Configurations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean error in centile scores across ComBat configurations")

subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
  ggplot(aes(x=sim.site, y=mean_cent_abs.diff, fill=sex)) +
  geom_violin(trim=FALSE) +
  facet_wrap(~dataset) +
  labs(y="Magnitude Centile Error", x = "Simulated Site") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean magnitude error in centile scores across \nComBat configurations")

#look at sex distributions of ppts across sites
subj_sum.df %>%
  ggplot() +
  geom_bar(aes(x=sim.site, fill=sex))+
  ggtitle("Distributions of male and female subjects across simulated sites")
```

Significance testing (w/in sub. mean abs centile error)  - using wilcox (mann whitney u) since distributions are skewed but variances are ~= across sexes

```{r}
sex.bias.wilcox.tests(subj_sum.df)
sex.bias.wilcox.tests(subj_sum.df, to_test="mean_cent_diff")
```

Keeping these analyses in case I want to use one of the above violin/boxplots that show w/in subject averages (across features). However, while these results I think are good to have from a "be aware of how biological variables are distributed across your study sites" perspective, it doesnt really say anything interesting froma "combat methods" perspective.

Testing for significant M:F differences in centile erros WITHIN each feature!

@@

```{r}
plt.df <- diffs.df %>%
  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("diff_", pheno_list)) %>% #select centile errors only
  dplyr::filter(dataset == "cf.gam" | dataset == "cf.gamlss") %>% #drop refA levels
  pivot_longer(cols=starts_with("diff_"), names_to="pheno", names_prefix="diff_", values_to = "diff") %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         diff = diff*100,
         pheno_cat = factor(case_when(
    pheno %in% vol_list_global ~ "Global Volume",
    pheno %in% vol_list_regions ~ "Regional Volume",
    pheno %in% sa_list ~ "Regional SA",
    pheno %in% ct_list ~ "Regional CT",
    TRUE ~ NA_character_), ordered=TRUE, levels=c("Global Volume", "Regional Volume", "Regional SA", "Regional CT"))) %>%
    mutate(pheno = factor(pheno, ordered=TRUE, levels=c(vol_list_global, vol_list_regions, sa_list, ct_list)))#sort w/in pheno_cat
```

stat test re-do w wilcox 
```{r}
stat.test <- plt.df %>%
  group_by(dataset, pheno) %>%
  wilcox_test(diff ~ sex) %>%
  adjust_pvalue(method = "fdr") %>%
  add_significance("p.adj")
```
```{r}
stat.test2 <- stat.test %>%
  add_xy_position(x = "pheno", dodge = 0.8, fun="max") %>%
  mutate(p.adj.signif = ifelse(p.adj.signif == "ns", "ns", "*")) #simplify significance labels
```

big_violin_plt <- plt.df %>%
  ggplot(aes(x=pheno, y=diff)) +
  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
  # geom_violin(aes(fill=sex, color=sex)) +
  theme_linedraw(base_size = 11) +
  labs(y="Centile Error", x = "Brain Feature") +
  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), panel.spacing.x = unit(.1, "lines"), panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 3, tip.length = 0, hide.ns = TRUE, remove.bracket = TRUE) +
  # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
  theme(plot.title = element_text(size=12), axis.ticks.x = element_blank(), axis.text.x = element_blank(), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
  labs(fill = "Sex", color="Sex")
  # ggtitle("Male and female centile errors")
big_violin_plt
ggsave("figs/big_sex_violin_plt.jpeg", big_violin_plt, width=12, height=6, units="in")
```

```{r, eval=FALSE}
big_violin_plt_wlabels <- plt.df %>%
  ggplot(aes(x=pheno, y=diff)) +
  geom_violin(aes(x=interaction(pheno,pheno_cat, sep = "&"), y=diff, fill=sex, color=sex)) +
  # geom_violin(aes(fill=sex, color=sex)) +
  theme_linedraw(base_size = 11) +
  labs(y="Centile Error", x = "Brain Feature") +
  theme(plot.title = element_text(size=12)) +
  stat_pvalue_manual(stat.test2, label = "p.adj.signif", label.size = 2.5, tip.length = 0, remove.bracket = TRUE) +
  # facet_grid2(dataset ~ pheno_cat, labeller=labeller(dataset=dataset_names), scales="free_x", space = "free_x", switch="x", strip = custom_strip) +
  facet_wrap(~dataset, labeller=labeller(dataset=dataset_names), nrow=2) +
  guides(x = ggh4x::guide_axis_nested(delim = "&")) +
  theme(plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust=0.4, size=6), ggh4x.axis.nesttext.x = element_text(angle = 90, vjust=0.4), legend.position="top") +
  labs(fill = "Sex", color="Sex")
  # ggtitle("Male and female centile errors")
big_violin_plt_wlabels
ggsave("figs/big_sex_violin_plt_wlabels.jpeg", big_violin_plt_wlabels, width=16, height=9, units="in")
```

@@

p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-biases in centile score errors across combat configurations", 
               color = "black", size = 14, hjust=.65))
p_sexannotate
ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
saveRDS(p_sexannotate, file="figs/sex_diff_win_feat.RDS")
```

Sex Effect is the median male centile error - median female centile error (i.e. units are in centiles). Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)? Honestly, not the most straightforward interpretation, i know - may make more sense to just plot the median errors separately in males and females than to plot the differences between the two.

look at # males vs females w extreme mean centiles

@@

@@

w/in feature sex-biases
```{r, message=FALSE, eval=FALSE}
noext.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(no_ext, feature_list=pheno_diff_list)
saveRDS(noext.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_cent_sexbias_all_noext_out.RDS")
```

@@

## Z-SCORES

@@

Foregoing anova based on emails with andrew (anova violates assumtptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)

Within each phenotype, iterate paired t test of ranks between each combo of combat dfs (nonparametric per discussion with Andrew). (see Zimmerman & Zumbo, 1993). Using custom functions from `ranked_welch_tests.R`.

@@

plot w ref sites
```{r, eval=FALSE}
#get means w/in feature 
mean.z.diffs.long <- diffs.df %>%
  dplyr::select(sex, dataset, pheno, z.pheno_abs.diff.list) %>%
  mutate(dataset = factor(dataset, levels = c("ukb_CN", "cf", "cf.lm", "cf.gam", "cf.gamlss", "cf.lm_refA", "cf.gam_refA", "cf.gamlss_refA"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  pivot_longer(cols=starts_with("abs.diff_"), values_to="z_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno=gsub(".z", "", pheno)) %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

mean.z.diffs.long %>%
  dplyr::filter(dataset != "ukb_CN") %>%
  ggplot() +
  geom_histogram(aes(x=z_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_grid(dataset~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  labs(y="Percent Features", x = "Mean Absolute Z-Score Error") +
  ggtitle("Absolute errors in zscores of brain structure features across ComBat processing pipelines")
```

@@

```{r}
gam_better.z <- gam.gamlss.diffs %>%
  group_by(dataset) %>%
  summarize_at(z.pheno_abs.diff.list, median) %>%
  ungroup() %>%
  pivot_longer(cols = all_of(z.pheno_abs.diff.list), names_to="pheno") %>%
  pivot_wider(names_from = dataset, values_from=value, names_glue = "med_{dataset}") %>%
  mutate(gam_minus_gamlss = (med_cf.gam - med_cf.gamlss),
         pheno = sub("abs.diff_", "", pheno)) %>%
  ungroup() %>%
  dplyr::filter(gam_minus_gamlss < 0)

gam_better.z$pheno

ns_gam_gamlss_comp.z <- full_z_res %>%
  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE)

intersect(gam_better.z$pheno, ns_gam_gamlss_comp.z$pheno)
```

### Sex-Biases

Visualize

```{r, eval=FALSE}
subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot() +
  geom_boxplot(aes(x=sex, y=mean_z_abs.diff, fill=sex)) +
  labs(y= "Magnitude Z Error", x = "ComBat Configurations") +
  facet_wrap(~dataset, scales="free_y") +
  theme(legend.position="none") +
  ggtitle("Within-subject mean magnitude of error in z-scores across ComBat configurations")

subj_sum.df %>%
  ggplot(aes(x=dataset, y=mean_z_diff, fill=sex)) +
  geom_violin(trim=FALSE, scale="width") +
  labs(y="Z Error", x = "ComBat Configurations") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean error in z-scores across ComBat configurations")

subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot(aes(x=sim.site, y=mean_z_abs.diff, fill=sex)) +
  geom_violin(trim=FALSE) +
  facet_wrap(~dataset) +
  labs(y="Magnitude Z Error", x = "Simulated Site") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  ggtitle("Within-subject mean magnitude error in z-scores across \nComBat configurations")

#look at sex distributions of ppts across sites
subj_sum.df %>%
  ggplot() +
  geom_bar(aes(x=sim.site, fill=sex), position="fill")+
  ggtitle("Distributions of male and female subjects across simulated sites")
```

@@

skewed but ~ variances? sticking with t test of ranks. Also calculating the median z-score errors of males and females

@@

@@

  df <- merge.df %>%
    dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, list) %>% #get correct cols
    group_by(participant) %>% 
    dplyr::filter(!any((get(pheno) < -2 | get(pheno) > 2) & dataset == "ukb_CN")) %>% #drop ppts if raw z score is < -2 or > 2
    ungroup() %>%
    dplyr::filter(dataset != "ukb_CN") #now drop raw vals
  z.df_list[count] <- list(df)
  count <- count + 1
}
length(z.df_list) == length(pheno_list.z)
```
merge into single dataframe
```{r, message=FALSE}
z.no_ext <- Reduce(full_join, z.df_list)

@@

w/in feature sex-biases
```{r, message=FALSE}
noext.z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(z.no_ext, feature_list=z.pheno_diff_list)
saveRDS(noext.z.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_z_sexbias_all_noext_out.RDS")
```

@@

## CENTILE FANS

@@

```{r, eval=FALSE}
global.cent.df <- centiles.df %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  dplyr::filter(!grepl('-refA', dataset))

for (feature in unique(global.cent.df$pheno)) {

  df <- global.cent.df %>%
    dplyr::filter(pheno == feature)

 plot <- ggplot() +
      #geom_point(aes(x=ukb_cn_df$age_days, y=ukb_cn_df[[feature]]), alpha=0.5) +
      #scale_colour_manual(values=c("#5AAE61FF", "#9970ABFF")) +
      geom_line(aes(x=df$age_days, y=df$M_centile_1, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_2, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_3, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_4, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_5, color=df$dataset), linewidth=1.5) +
      geom_line(aes(x=df$age_days, y=df$M_centile_6, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_7, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_8, color=df$dataset)) +
      geom_line(aes(x=df$age_days, y=df$M_centile_9, color=df$dataset)) +
      labs(title=paste("Male", feature, "brain chart"))

 print(plot)

>>>>

Main analyses of ComBatLS paper.

Relies on outputs derived/functions from:

* `./cubic_scripts/R_scripts/stats_tests.R`
1. `ukb_curation-eda.Rmd`
2. `qsub_combat.sh` -> `/cubic_scripts/R_scripts/combat_apply.R`
3. `qsub_basic_gamlss.sh` -> `/cubic_scripts/R_scripts/fit_basic_mod.R`
4. `qsub_centiles.sh` -> `/cubic_scripts/R_scripts/fit_centiles.R`
5. `gamlss_parse.sh` -> `/cubic_scripts/R_scripts/gamlss_parser.R`

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE)
if(!require('pacman')) {
  install.packages('pacman')
}
pacman::p_load(data.table, ggplot2, tidyverse, ggseg, ggsegExtra, paletteer, ggpubr, gridExtra, lemon, parallel, rstatix, lme4, interactions, ggrepel, ggrain, ggh4x)
```
```{r}
source("./cubic_scripts/R_scripts/stats_tests.R")
```

@@

# Load Data

@@

QC: check for ppts with centiles ~ 1 or 0 and make sure Z scores exist

```{r}
combined_preds %>% 
  filter_at(vars(pheno_list), any_vars(. %in% c(1, 0))) %>%
  dplyr::select(participant) %>%
  unique() %>%
  nrow()

combined_preds %>% 
  filter_at(vars(pheno_list.z), any_vars(. %in% c(NA, Inf, -Inf))) %>%
  dplyr::select(participant) %>%
  unique() %>%
  nrow()
```

saving out csv forces some centile scores to round up to 1, but z scores are still being calculated correctly.

@@

# CENTILES

Do centile scores change based on combat version?
Foregoing anova based on emails with Andrew (anova violates assumptions of no equal variance but there's no option for kruskall-wallace with paired samples - which these are, since it's the same subjects)

@@

```{r, eval=FALSE}
full_cent_res <- centile.t.tests.full_result(diffs.df, feature_list=pheno_abs.diff.list)
saveRDS(full_cent_res, "./derivatives/ukb_basic/featurewise_cent_t_tests_all_out.RDS") #running once and saving because it takes a long time
```

@@

bigger group -> relatively worse fit

@@

get stat results for global phenos:
```{r, message=FALSE}
full_cent_res %>%
  dplyr::filter(pheno %in% vol_list_global) %>%
  dplyr::filter(!grepl('_refA', comp)) %>%
  summarise(n_sig = sum(sig_fdr),
            nrow = nrow(.))

#n occurances of combatLS having significantly larger errors
full_cent_res %>%
  dplyr::filter(bigger_group == "cf.gamlss")%>%
  dplyr::filter(!grepl('_refA', comp)) %>%
  summarise(n_sig = sum(sig_fdr),
            nrow = nrow(.))
```

global_abs.diffs <- diffs.df %>%
  dplyr::select(participant, sex, age_days, sim.site, dataset, paste0("abs.diff_", vol_list_global)) %>%
  pivot_longer(cols=starts_with("abs.diff_"), names_to="pheno", names_prefix="abs.diff_", values_to = "abs.diff")

@@

```{r}
ns_gam_gamlss_comp <- full_cent_res %>%
  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE & bigger_group == "cf.gamlss")
nrow(ns_gam_gamlss_comp)
```

@@

### Sex-Differences

Testing for significant M:F differences in centile errors within each feature

@@

@@

p_sexannotate <- annotate_figure(p_sex, top = text_grob("Figure 2A. Sex-differences in centile score errors across combat configurations", 
               color = "black", size = 14, hjust=.65))
p_sexannotate
ggsave("figs/sex_diff_win_feat.jpeg", p_sexannotate, width=7, height=8)
saveRDS(p_sexannotate, file="figs/sex_diff_win_feat.RDS")
```

Sex Effect is the median male centile error - median female centile error (i.e. units are in centiles). Therefore, + sex effect (blue) means female scores tend to have more positive errors (TRUE centile > calculated centile) and - sex effect (green) means females have more negative errors (TRUE centile < calculated centile)

@@

bigger group -> relatively worse fit

@@

w/in feature sex-differences
```{r, message=FALSE, eval=FALSE}
noext.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(no_ext, feature_list=pheno_diff_list)
saveRDS(noext.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_cent_sexbias_all_noext_out.RDS")
```

@@

# Z-SCORES

@@

Within each phenotype, iterate paired t test of ranks between each combo of combat dfs using custom functions from `ranked_welch_tests.R`.

@@

@@

```{r}
ns_gam_gamlss_comp.z <- full_z_res %>%
  dplyr::filter(comp=="cf.gamlss_cf.gam" & sig_fdr == TRUE & bigger_group == "cf.gamlss")
nrow(ns_gam_gamlss_comp.z)
```

### Sex-Differences

@@

skewed but ~ variances, sticking with t test of ranks. Also calculating the median z-score errors of males and females

@@

Get raw z-scores and merge with z-score differences

@@

  df <- merge.df %>%
    dplyr::select(participant, sex, age_days, sim.site, Source_File, dataset, list) %>% #get correct cols
    group_by(participant) %>% 
    dplyr::filter(!any((get(pheno) < -2 | get(pheno) > 2) & dataset == "ukb_CN")) %>% #drop ppts if raw z score is < -2 or > 2
    ungroup() %>%
    dplyr::filter(dataset != "ukb_CN") #now drop raw vals
  z.df_list[count] <- list(df)
  count <- count + 1
}
length(z.df_list) == length(pheno_list.z)
```

merge into single dataframe
```{r, message=FALSE}
z.no_ext <- Reduce(full_join, z.df_list)

@@

w/in feature sex-differences
```{r, message=FALSE}
noext.z.sex_t_tests_in_feat.df <- sex.bias.feat.t.tests(z.no_ext, feature_list=z.pheno_diff_list)
saveRDS(noext.z.sex_t_tests_in_feat.df, "./derivatives/ukb_basic/featurewise_z_sexbias_all_noext_out.RDS")
```

@@

# CENTILE FANS

@@
